---
title: "PROJECT 406"
output: pdf_document
date: "2024-10-23"
---

#



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages('tibble')
library(tibble)
library(tidyr)
library(dplyr)
library(ggplot2)
```

```{r setup, include=FALSE}
fantasy_clean <- read.csv('fantasy_cleaned.csv')
fantasy_cleaned1 <- as_tibble(fantasy_clean, .name_repair = "unique")
fantasy_cleaned1$PPR <- as.numeric(fantasy_cleaned1$PPR)
```

```{r}
df_filtered <- fantasy_cleaned1 %>% filter(G > 7)
print(df_filtered)
```

##Based on Notes
#Calculate year-to-year variance in PPR by position
```{r}
df_filtered_grouped <- fantasy_cleaned1 %>%
  group_by(Player, FantPos, Year) %>%
  summarise(PPR = sum(PPR, na.rm = TRUE)) %>%
  ungroup()
```


```{r}
df_pivot <- tidyr::pivot_wider(df_filtered_grouped, 
                               names_from = Year, 
                               values_from = PPR)

variance_by_position <- df_pivot %>%
  group_by(FantPos) %>%
  summarise(across(everything(), ~ var(.x, na.rm = TRUE)))
print(variance_by_position)
```

#Bootstrap to estimate variance for WRs vs other positions
```{r}
bootstrap_variance <- function(data, n_bootstrap = 1000) {
  variances <- numeric(n_bootstrap)
  for (i in 1:n_bootstrap) {
    sample_data <- sample(data, replace = TRUE)
    variances[i] <- var(sample_data, na.rm = TRUE)
  }
  return(variances)
}
```


```{r}
wr_data <- df_pivot %>% filter(FantPos == "WR") %>% select(-Player, -FantPos)
wr_variance <- bootstrap_variance(unlist(wr_data))
rb_data <- df_pivot %>% filter(FantPos == "RB") %>% select(-Player, -FantPos)
rb_variance <- bootstrap_variance(unlist(rb_data))

print(mean(wr_variance))
print(mean(rb_variance))
```

#Exclude top 24 players and recalculate
```{r}
df_mean_ppr <- df_filtered_grouped %>%
  group_by(Player, FantPos) %>%
  summarise(mean_PPR = mean(PPR, na.rm = TRUE)) %>%
  ungroup()

top_24_players <- df_mean_ppr %>%
  group_by(FantPos) %>%
  top_n(24, mean_PPR)

df_filtered_excluded <- df_filtered_grouped %>%
  filter(!(Player %in% top_24_players$Player))

df_pivot_excluded <- tidyr::pivot_wider(df_filtered_excluded, 
                                        names_from = Year, 
                                        values_from = PPR)

variance_by_position_excluded <- df_pivot_excluded %>%
  group_by(FantPos) %>%
  summarise(across(everything(), ~ var(.x, na.rm = TRUE)))

print(variance_by_position_excluded)
```

#Coefficient of Variation by Position
```{r}
cv_by_position <- fantasy_cleaned1 %>%
  group_by(FantPos) %>%
  summarise(mean_FantPt = mean(as.numeric(PPR), na.rm = TRUE),
            sd_FantPt = sd(as.numeric(PPR), na.rm = TRUE)) %>%
  mutate(CV = sd_FantPt / mean_FantPt)

print(cv_by_position)
```

#Monte Carlo Simulation for Draft Strategies
```{r}
set.seed(123)  

simulate_draft_strategy <- function(n_sim, positions_order) {
  results <- numeric(n_sim)
  
  for (i in 1:n_sim) {
    draft <- fantasy_cleaned1 %>%
      filter(FantPos %in% positions_order) %>%
      group_by(Player, FantPos) %>%
      summarise(total_PPR = sum(as.numeric(PPR), na.rm = TRUE)) %>%
      ungroup() %>%
      sample_n(length(positions_order)) %>%
      pull(total_PPR)
    
    results[i] <- sum(draft)
  }
  
  return(results)
}

# Not working
# wr_rb_strategy <- simulate_draft_strategy(n_sim = 1000, positions_order = c("WR", "RB"))
# rb_wr_strategy <- simulate_draft_strategy(n_sim = 1000, positions_order = c("RB", "WR"))

# mean(wr_rb_strategy)
# mean(rb_wr_strategy)
```


#Bootstrapping for Hypothesis Testing on CV
```{r}
bootstrap_cv <- function(data, n_bootstrap = 1000) {
  boot_cvs <- numeric(n_bootstrap)
  
  for (i in 1:n_bootstrap) {
    sample_data <- data %>% sample_frac(replace = TRUE)
    mean_PPR <- mean(sample_data$PPR, na.rm = TRUE)
    sd_PPR <- sd(sample_data$PPR, na.rm = TRUE)
    boot_cvs[i] <- sd_PPR / mean_PPR
  }
  
  return(boot_cvs)
}

wr_data <- fantasy_cleaned1 %>% filter(FantPos == "WR")
wr_bootstrap <- bootstrap_cv(wr_data)

rb_data <- fantasy_cleaned1 %>% filter(FantPos == "RB")
rb_bootstrap <- bootstrap_cv(rb_data)

qb_data <- fantasy_cleaned1 %>% filter(FantPos == "QB")
rb_bootstrap <- bootstrap_cv(qb_data)

te_data <- fantasy_cleaned1 %>% filter(FantPos == "TE")
te_bootstrap <- bootstrap_cv(te_data)

t.test(wr_bootstrap, rb_bootstrap)
```

# Non-parametric Bootstrapping to Estimate Sampling Distributions
```{r}
bootstrap_sampling_distribution <- function(data, n_bootstrap = 1000) {
  boot_samples <- replicate(n_bootstrap, {
    sample_data <- data %>% sample_frac(replace = TRUE)
    mean_PPR <- mean(sample_data$PPR, na.rm = TRUE)
    sd_PPR <- sd(sample_data$PPR, na.rm = TRUE)
    return(sd_PPR / mean_PPR)
  })
  
  return(boot_samples)
}

wr_bootstrap_dist <- bootstrap_sampling_distribution(wr_data)
rb_bootstrap_dist <- bootstrap_sampling_distribution(rb_data)
qb_bootstrap_dist <- bootstrap_sampling_distribution(qb_data)
te_bootstrap_dist <- bootstrap_sampling_distribution(te_data)

```

```{r}
# COLOR CODING
position_colors <- c("QB" = "red", "RB" = "blue", "WR" = "purple", "TE" = "green")
```

```{r}
boot_data <- data.frame(
  CV = c(qb_bootstrap_dist, rb_bootstrap_dist, wr_bootstrap_dist, te_bootstrap_dist),
  Position = factor(rep(c("QB", "RB", "WR", "TE"), each = length(wr_bootstrap_dist)))
)

ggplot(boot_data, aes(x = CV, fill = Position)) + 
  geom_density(alpha = 0.5) + 
  labs(title = "Bootstrapped CV Distribution by Position", 
       x = "Coefficient of Variation", 
       y = "Density") +
  xlim(.9, 1.15) + 
  theme_minimal() + 
  scale_fill_brewer(palette = "Set1")
```

##Personnel 
```{r}
#Player Consistency Analysis with Time Series or Distribution                             
player_consistency <- fantasy_cleaned1 %>%
  filter(G > 8, FantPos != 'FB') %>% 
  select(Player, FantPos, PPR) %>%
  drop_na()


ggplot(player_consistency, aes(x = PPR, fill = FantPos)) +
  geom_density(alpha = 0.5) +
  labs(title = "Distribution of PPR Point Scored by Position",
       x = "PPR Fantasy Points",
       y = "Density") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") 

ggplot(player_consistency, aes(x = PPR, fill = FantPos)) + 
  geom_density(alpha = 0.5) + 
  labs(title = "Individual Distribution of PPR Points Scored",
       x = "PPR Fantasy Points", 
       y = "Density") + 
  theme_minimal() + 
  scale_fill_brewer(palette = "Set1") +
  facet_wrap(~ FantPos, ncol = 2)

```

```{r}
library(dplyr)
library(gt)

# Calculate summary statistics for each position
summary_table <- player_consistency %>%
  group_by(Pos = FantPos) %>%
  summarize(
    Min = min(PPR, na.rm = TRUE),
    Q1 = quantile(PPR, 0.25, na.rm = TRUE),
    Median = median(PPR, na.rm = TRUE),
    Mean = mean(PPR, na.rm = TRUE),
    Q3 = quantile(PPR, 0.75, na.rm = TRUE),
    Max = max(PPR, na.rm = TRUE)
  )

# Display as a styled table with rounded values
summary_table %>%
  gt() %>%
  tab_header(title = "Summary Statistics for PPR by Position") %>%
  fmt_number(columns = vars(Min, Q1, Median, Mean, Q3, Max), decimals = 2)


```

```{r}
library(dplyr)
library(knitr)

# Calculate and round summary statistics for each position
summary_table <- player_consistency %>%
  group_by(Pos = FantPos) %>%
  summarize(
    Min = round(min(PPR, na.rm = TRUE), 2),
    Q1 = round(quantile(PPR, 0.25, na.rm = TRUE), 2),
    Median = round(median(PPR, na.rm = TRUE), 2),
    Mean = round(mean(PPR, na.rm = TRUE), 2),
    Q3 = round(quantile(PPR, 0.75, na.rm = TRUE), 2),
    Max = round(max(PPR, na.rm = TRUE), 2)
  )

# Display as a nicely formatted table
kable(summary_table, caption = "Summary Statistics for PPR by Position")

```



```{r}
player_consistency_top1000 %>%
  arrange(-PPR)
```



```{r}
player_consistency_top1000 <- player_consistency %>%
  group_by(FantPos) %>%
  arrange(desc(PPR)) %>%
  slice_head(n = 1000) %>%
  ungroup() 

ggplot(player_consistency_top1000, aes(x = PPR, fill = FantPos)) + 
  geom_density(alpha = 0.5) + 
  labs(title = "Distribution of PPR Points Scored by Position",
       x = "PPR Fantasy Points", 
       y = "Density") + 
  theme_minimal() + 
  scale_fill_brewer(palette = "Set1") 
```
